Limits to Ecological Forecasting: Estimating Uncertainty for Critical Transitions with Deep Learning

### 2022-09-20 manuscript/manuscript.Rmd
fix var names in dummy 'predicted' entry

```{r figure1, fig.cap="Forecast scenarios. A. The Hopf bifurcation: a stable node develops into a limit cycle which gradually grows larger in this predator-prey model. B. The saddle-node bifurcation in a single species. C. The stochastic transition in a single species.  Plots show historical data used to train the algorithm in purple, and replicate simulations of the true dynamics ('future ensemble') in yellow. Note how the characteristic time for the critical transition varies across the transitions. We will examine forecasts of various models (Fig 2) which will each produce probablistic forecast distributions (blue, Fig 3-5) seeking to match the true future ensemble (yellow) as closely as possible.", fig.height=6}
hopf <- read_csv("../greta/forecasts/simulation=hopf/hopf_increasing.csv.gz")
panelA <- hopf |> 
  mutate(type = as.factor(type)) |>
  filter(type %in% c("historical", "true"), i < 10) |>
  rename(species = variable) |> 
  mutate(species = forcats::fct_recode(species, host = "H", parasitoid = "P")) |>
  bind_rows(tibble(i=1,t=1,type="predicted", variable="H", value=NA)) |> #placeholder
  mutate(type = forcats::fct_recode(type, "future ensemble" = "true", "forecast" = "predicted")) |>
  ggplot(aes(t, value, group=interaction(species, i, type), col=type, lty=species)) + 
  geom_line(show.legend = TRUE, alpha = 0.7) +
  theme_bw() +  
  scale_color_viridis_d(guide="none") +
  ggtitle("A. Hopf bifurcation")+
  theme(legend.position="top")

>>>>

```{r figure1, fig.cap="Forecast scenarios. A. The Hopf bifurcation: a stable node develops into a limit cycle which gradually grows larger in this predator-prey model. B. The saddle-node bifurcation in a single species. C. The stochastic transition in a single species.  Plots show historical data used to train the algorithm in purple, and replicate simulations of the true dynamics ('future ensemble') in yellow. Note how the characteristic time for the critical transition varies across the transitions. We will examine forecasts of various models (Fig 2) which will each produce probablistic forecast distributions (blue, Fig 3-5) seeking to match the true future ensemble (yellow) as closely as possible.", fig.height=6}
hopf <- read_csv("../greta/forecasts/simulation=hopf/hopf_increasing.csv.gz")
panelA <- hopf |> 
  mutate(type = as.factor(type)) |>
  filter(type %in% c("historical", "true"), i < 10) |>
  rename(species = variable) |> 
  mutate(species = forcats::fct_recode(species, host = "H", parasitoid = "P")) |>
  bind_rows(tibble(i=1,t=1,type="predicted", species="host", value=NA)) |> #placeholder
  mutate(type = forcats::fct_recode(type, "future ensemble" = "true", "forecast" = "predicted")) |>
  ggplot(aes(t, value, group=interaction(species, i, type), col=type, lty=species)) + 
  geom_line(alpha = 0.7) +
  theme_bw() +  
  scale_color_viridis_d(guide="none") +
  ggtitle("A. Hopf bifurcation")+
  theme(legend.position="top")
